\section{Overview}

This chapter summarizes how Code Guardian was implemented as a local-first security assistant for Visual Studio Code. The focus is on deployable engineering decisions: local inference boundaries, deterministic output handling, latency guardrails, and developer-controlled remediation.

The implementation follows the concept from Chapter~\ref{chap:concept} while emphasizing practical concerns that affect reproducibility and day-to-day usability. The chapter first introduces the technical stack and runtime boundary, then explains the detection pipeline, the main components, responsiveness mechanisms, user-facing integration, and repair safety.

\section{Technology Stack and Runtime Boundary}

Code Guardian is implemented as a VS Code extension with a local Node.js backend and local Ollama inference. Retrieval augmentation, when enabled, uses a local embedding model and local vector store.

\begin{itemize}
  \item \textbf{IDE integration:} VS Code extension APIs for diagnostics, code actions, commands, and WebViews.
  \item \textbf{Analysis backend:} Node.js service that builds prompts, calls models, validates output, and returns normalized findings.
  \item \textbf{Inference runtime:} Ollama on localhost; no cloud model endpoint is required.
  \item \textbf{RAG layer:} local embedding + local vector index over curated CWE/OWASP/CVE guidance.
\end{itemize}

The evaluated deployment keeps source code, prompts, and findings on-device during analysis. Optional knowledge refresh accesses only public vulnerability metadata and can be disabled for offline use.

\section{End-to-End Detection Pipeline}

The implementation follows a fixed pipeline to improve reproducibility and reduce integration risk:
\begin{enumerate}
  \item Extract analysis scope (selection, function, file, or workspace unit).
  \item Build model prompt (LLM-only or LLM+RAG).
  \item Enforce JSON-structured response contract.
  \item Parse and normalize findings.
  \item Map findings to IDE diagnostics and optional quick fixes.
\end{enumerate}

\subsection*{Scope extraction}
Function-level extraction is the default for interactive usage because it limits latency and token usage. Broader scopes (file/workspace) are available for explicit audits.

\subsection*{Prompt construction}
LLM-only mode prioritizes lower latency. RAG mode injects retrieved security context to improve grounding when model behavior supports it.

\subsection*{Structured output handling}
The analyzer enforces a strict JSON contract and applies defensive parsing to tolerate minor output noise. Parse failures degrade safely to empty findings or warning paths instead of crashing the extension workflow.

\section{Core Components}

\subsection*{Analyzer}
The analyzer orchestrates prompt assembly, model invocation, timeout handling, and response validation. It converts model output into a stable schema used by downstream diagnostics and evaluation scripts.

\subsection*{Diagnostics mapper}
Findings are mapped to VS Code diagnostics with severity and location metadata. The mapper tolerates imperfect line localization by clamping ranges to valid document boundaries.

\subsection*{RAG manager}
The retrieval manager handles local index queries and retrieval filtering. It supports updating the local knowledge base from public sources while preserving the local source-code boundary.

\subsection*{Workspace scanner}
Batch scanning processes files under configurable size/scope limits and aggregates results for dashboard-style review.

\section{Orchestration and Responsiveness}

Interactive IDE usage requires explicit guardrails:
\begin{itemize}
  \item \textbf{Debouncing:} analysis is delayed after edits to avoid redundant invocations.
  \item \textbf{Scope limits:} large functions/files are skipped or deferred to audit mode.
  \item \textbf{Caching:} repeated analyses use TTL-based cache entries to reduce local inference overhead.
  \item \textbf{Mode selection:} faster modes are favored during active editing; stronger modes are used for explicit scans.
\end{itemize}

These mechanisms reduce interruption cost and make local inference practical on developer hardware, while preserving deterministic failure handling.

\section{User-Facing Integration}

Security feedback is surfaced in native IDE flows:
\begin{itemize}
  \item inline diagnostics and hover explanations,
  \item optional quick-fix actions for model-proposed repairs,
  \item command-driven scans and model/mode controls,
  \item dashboard-oriented workspace summaries.
\end{itemize}

The design principle is assistive, not autonomous: findings and fixes are review artifacts under developer control.

\section{Repair Safety}
\label{sec:impl-repair-safety}

Repair generation is intentionally conservative:
\begin{itemize}
  \item fixes are suggested as diffs, not silently applied;
  \item users decide whether to apply or reject each change;
  \item no automatic commit/deploy path is implemented;
  \item functional/security re-validation remains a developer responsibility.
\end{itemize}

In practice, repairs are useful when changes are local and pattern-based (for example parameterization or sanitizer insertion). Complex cases still require domain context and manual adjustment.

\section{Implementation Summary}

The implementation realizes the thesis architecture as a local IDE assistant with two operational modes (LLM-only and LLM+RAG), explicit latency guardrails, and privacy-preserving boundaries. These concrete mechanisms provide the basis for interpreting the evaluation results in Chapter~\ref{chap:evaluation}.
