\section{Qualitative Case Studies and Repair Suggestions}
\label{sec:eval-case-studies}

Quantitative metrics summarize overall detection behavior, but IDE usefulness also depends on explanation clarity and repair quality. This section reports qualitative observations from representative cases in the recorded ablation run.

\subsection*{Representative Cases}

\textbf{Case A (True Positive with direct remediation): SQL injection.}
Sample ID: \texttt{sql-injection-1}.
In the best-performing configuration (\texttt{qwen3:8b} with RAG), the model flagged direct string concatenation in the query and suggested parameterized queries. The suggested remediation aligned with the expected fix and was minimal.

\textbf{Case B (False Positive on secure sample): safe process invocation.}
Sample ID: secure snippet with \texttt{execFile} array arguments.
In \texttt{qwen3:8b} (LLM+RAG), this secure sample was still flagged, with a broad path-sanitization recommendation. This illustrates residual over-warning even in the strongest configuration.

\textbf{Case C (False Negative): NoSQL injection.}
Sample ID: NoSQL injection representative case.
In \texttt{qwen3:8b} (LLM+RAG), this vulnerable sample was missed in a representative run and no remediation suggestion was produced.

\textbf{Case D (True Positive): cross-site scripting via \texttt{innerHTML}.}
Sample ID: XSS representative case.
In \texttt{qwen3:8b} (LLM+RAG), the model correctly flagged unsafe HTML sink usage and proposed replacing \texttt{innerHTML} with safer rendering behavior.

\subsection*{Repair Suggestion Quality (R4)}

Repair suggestions are evaluated qualitatively against three criteria:
\begin{itemize}
  \item \textbf{Security adequacy:} does the fix mitigate the vulnerability class (e.g., parameterization for SQL injection)?
  \item \textbf{Minimality:} does it avoid unnecessary refactoring?
  \item \textbf{Applicability:} does it fit the code context and available APIs?
\end{itemize}

Because Code Guardian operates inside the IDE, repair suggestions are intentionally presented as \emph{optional} quick fixes, enabling developer review and preventing silent modifications.

\subsection*{Quantitative Repair Analysis}

To provide quantitative evidence for R4 compliance, we manually analyzed 25 repair suggestions sampled from the \texttt{qwen3:8b} (LLM+RAG) configuration across diverse vulnerability types. Each repair was evaluated against four criteria: (1) whether a fix was provided, (2) whether the fix was semantically correct, (3) whether it aligned with the expected remediation strategy, and (4) whether it included executable code rather than only prose guidance.

\begin{table}[H]
  \centering
  \caption{Manual repair quality assessment ($n=25$ sampled repairs from \texttt{qwen3:8b} with RAG).}
  \label{tab:repair-quality}
  \footnotesize
  \renewcommand{\arraystretch}{1.15}
  \begin{tabular}{p{0.38\textwidth}ccp{0.28\textwidth}}
    \toprule
    Criterion & Count & Percentage & Notes \\
    \midrule
    Fix provided (non-null) & 19/25 & 76.0\% & 6 cases returned no suggested fix \\
    Semantically correct & 17/19 & 89.5\% & Of those with fixes provided \\
    Aligned with expected fix & 18/19 & 94.7\% & Addressed same vulnerability class \\
    Contains executable code & 8/19 & 42.1\% & Others were prose guidance only \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Breakdown by vulnerability type.}
Table~\ref{tab:repair-by-type} reports repair provision rates across the vulnerability categories represented in the sample.

\begin{table}[H]
  \centering
  \caption{Repair provision rate by vulnerability type ($n=25$ sampled cases).}
  \label{tab:repair-by-type}
  \footnotesize
  \begin{tabular}{lcc}
    \toprule
    Vulnerability Type & Cases & Fix Provided \\
    \midrule
    Command Injection & 3 & 3/3 (100\%) \\
    SQL Injection & 2 & 2/2 (100\%) \\
    Path Traversal & 3 & 3/3 (100\%) \\
    NoSQL Injection & 3 & 2/3 (67\%) \\
    LDAP Injection & 3 & 2/3 (67\%) \\
    Insecure Deserialization & 3 & 3/3 (100\%) \\
    Authentication Bypass & 3 & 2/3 (67\%) \\
    Weak Cryptography & 3 & 2/3 (67\%) \\
    Prototype Pollution & 3 & 0/3 (0\%) \\
    Race Condition & 1 & 0/1 (0\%) \\
    \midrule
    \textbf{Total} & \textbf{25} & \textbf{19/25 (76\%)} \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
The repair analysis reveals three patterns. First, common vulnerability classes (SQL injection, command injection, path traversal) consistently receive appropriate fix suggestions, typically recommending parameterization, safe APIs, or input validation. Second, more complex patterns (prototype pollution, race conditions) often fail to generate any repair suggestion, indicating model limitations for nuanced vulnerability classes. Third, most repairs are prose-based guidance (``use parameterized queries'') rather than directly applicable code patches; while this is useful for developer education, it increases the effort required to apply fixes.

These findings support treating Code Guardian repairs as \emph{assistive suggestions} requiring developer judgment, consistent with the R4 requirement for actionable but not autonomous remediation.

\subsection*{Explanation Quality Analysis (R3)}

To provide quantitative evidence for R3 (Explainability), we analyzed the same 25 sampled detections from \texttt{qwen3:8b} (LLM+RAG) for explanation quality indicators. Each explanation (the \texttt{message} field in structured output) was evaluated against three criteria reflecting IDE usability.

\begin{table}[H]
  \centering
  \caption{Explanation quality assessment ($n=25$ sampled detections from \texttt{qwen3:8b} with RAG).}
  \label{tab:explanation-quality}
  \footnotesize
  \renewcommand{\arraystretch}{1.15}
  \begin{tabular}{p{0.40\textwidth}ccp{0.24\textwidth}}
    \toprule
    Criterion & Count & Percentage & Notes \\
    \midrule
    References vulnerability type explicitly & 23/25 & 92.0\% & e.g., ``SQL injection'', ``path traversal'' \\
    Explains attack vector or risk & 19/25 & 76.0\% & e.g., ``allows attacker to inject...'' \\
    Mentions specific code element & 21/25 & 84.0\% & e.g., function name, variable, API \\
    Explanation length $>$ 50 characters & 24/25 & 96.0\% & Sufficient detail for context \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
The explanation quality analysis shows that the best-performing configuration produces diagnostically useful messages in most cases. Over 90\% of explanations explicitly name the vulnerability type, enabling developers to quickly categorize findings. Three-quarters explain the attack vector, providing actionable risk context. The high rate of code-element references (84\%) supports traceability between the warning and specific code locations. These indicators suggest that R3 is substantially met for the \texttt{qwen3:8b} configuration, though explanation quality varies across models and may degrade for weaker configurations.

\subsection*{Observed Failure Modes}

Observed failure modes in the run include:
\begin{itemize}
  \item \textbf{Over-warning:} many secure patterns are still flagged as vulnerable in multiple configurations.
  \item \textbf{Model-dependent recall shifts under RAG:} some models improve (notably \texttt{qwen3} variants), while others degrade.
  \item \textbf{Conservative under-warning:} low-latency configurations can miss many vulnerable samples.
  \item \textbf{High-latency audit modes:} strongest F1 configurations may still be too slow for inline workflows.
\end{itemize}

\subsection*{Error Taxonomy from Run Logs}

\begin{table}[H]
  \centering
  \caption{Observed error taxonomy in the ablation run.}
  \label{tab:eval-error-taxonomy}
  \footnotesize
  \renewcommand{\arraystretch}{1.15}
  \begin{tabular}{p{0.28\textwidth}p{0.30\textwidth}p{0.30\textwidth}}
    \toprule
    Error pattern & Representative evidence & Quantitative impact in this run \\
    \midrule
    Persistent secure-sample over-warning & \texttt{gemma3:4b}, \texttt{qwen3:4b}, \texttt{CodeLlama:latest} (both modes) & FPR remained 100\% in merged evaluation (all secure samples were flagged in these modes). \\
    Model-dependent RAG effect & \texttt{qwen3} vs.\ \texttt{gemma3:4b}/\texttt{CodeLlama} & RAG improved \texttt{qwen3:8b} (F1 58.12\% $\rightarrow$ 63.76\%) and \texttt{qwen3:4b} (54.90\% $\rightarrow$ 58.67\%), but reduced \texttt{gemma3:4b} and \texttt{CodeLlama:latest}. \\
    Conservative under-detection & \texttt{gemma3:1b} (LLM-only) & Recall remained comparatively low at 25.66\% (87/339) despite fast latency. \\
    High-latency high-F1 mode & \texttt{qwen3:8b} (LLM+RAG) & Best F1 (63.76\%) and best LLM FPR (26.67\%) came with higher latency (median 1524\,ms; mean 1827\,ms). \\
    \bottomrule
  \end{tabular}
\end{table}

These observations motivate the future work directions summarized in Chapter~\ref{chap:conclusion}.

\paragraph{Section implication.}
The qualitative cases complement the quantitative results by showing how metrics translate into developer experience. They confirm that the strongest configuration is useful because it often produces locally actionable explanations and repairs, but they also show why false positives and missed cases remain a serious practical concern. This supports the thesis position that Code Guardian should assist developer judgment rather than automate final security decisions.
