\section{Evaluation Metrics}
\label{sec:eval-metrics}

We evaluate \textit{Invox} along complementary axes: (i) semantic accuracy under paraphrase and set-size mismatch, (ii) behavior under typed schema constraints, and (iii) operational efficiency. 
Because MUC-4 slots often admit paraphrases or near-synonyms, we use embedding-based measurements for text and list fields, while keeping exact comparison for typed slots (enums, dates).
To disentangle empty-field effects, we report results in two regimes:
\emph{baseline} (standard scoring, where gold-empty $\wedge$ pred-empty contributes positively) and 
\emph{gold-nonempty-only} (scores computed \emph{only} when the gold slot is nonempty).

\subsection{Embedding-Based Accuracy (R2)}
\label{subsec:emb-acc}

For text and list-valued slots, we compare unordered sets of fillers using sentence embeddings (\texttt{all-MiniLM-L6-v2}) and cosine similarity. 
Let $G=\{g_i\}$ and $P=\{p_j\}$ be the gold and predicted sets. 
We define gold-anchored recall and prediction-anchored precision surrogates:
\begin{align}
\text{SoftCoverage}   &= \frac{1}{|G|}\sum_{i}\max_{j}\ \cos\!\big(\mathrm{enc}(g_i),\,\mathrm{enc}(p_j)\big), \\
\text{SoftSpecificity} &= \frac{1}{|P|}\sum_{j}\max_{i}\ \cos\!\big(\mathrm{enc}(p_j),\,\mathrm{enc}(g_i)\big).
\end{align}

Their symmetric mean gives a threshold-free Chamfer similarity,
\[
\text{Chamfer} \;=\; \tfrac{1}{2}\big(\text{SoftCoverage}+\text{SoftSpecificity}\big),
\]
and we also report a \emph{Soft-F1} (\textbf{SF1}) as the harmonic mean of SoftCoverage and SoftSpecificity:
\[
\text{SF1} \;=\; \frac{2\,\text{SoftCoverage}\cdot\text{SoftSpecificity}}{\text{SoftCoverage}+\text{SoftSpecificity}}.
\]
For \textbf{enums} and \textbf{dates}, we use \emph{exact} scoring (match = 1, otherwise 0), aligning with the typed nature of these slots. 
Free-form \textbf{locations} are treated as text (singletons) and scored via embeddings.

\paragraph{Two scoring regimes.} 
We compute all scores twice: (i) \textbf{baseline} (standard), and (ii) \textbf{gold-nonempty-only} (ignores any (doc,slot) where the gold is empty). 
This separation isolates true content quality from empty-field advantages.

\subsection{Decision Quality Under Typed Constraints (R1)}
\label{subsec:consistency}

Beyond similarity, we evaluate the model's \emph{fill decision} with a presence matrix over four outcomes per slot: 
gold-empty/pred-empty, gold-empty/pred-nonempty, gold-nonempty/pred-empty, gold-nonempty/pred-nonempty. 
From this we derive compact decision metrics:
\begin{itemize}
  \item \textbf{FDA} (Fill-Decision Accuracy): fraction of cases where the empty vs.\ nonempty choice matches gold.
  \item \textbf{HR} (Hallucination Rate): $\Pr(\text{pred nonempty} \mid \text{gold empty})$.
  \item \textbf{MR} (Missing Rate): $\Pr(\text{pred empty} \mid \text{gold nonempty})$.
  \item \textbf{RFA} (Required-Fill Accuracy): average slot score \emph{conditioned} on gold-nonempty \emph{and} pred-nonempty, capturing content quality when the model commits to a fill.
\end{itemize}
To summarize empty-field sensitivity we report the \textbf{Empty Advantage Index (EAI)}:
\[
\text{EAI} \;=\; \text{OBS} - \text{NES},
\]
where \textbf{OBS} is the overall baseline score and \textbf{NES} is the overall score in the gold-nonempty-only regime. Higher EAI indicates stronger reliance on empty$\leftrightarrow$empty matches.


\subsection{Consistency (R1-core): Two-Score Summary}
\label{subsec:r1-consistency-core}

\paragraph{Overall formatting.}
A record passes a slot’s lints if type/enums/regex (dates), and list hygiene hold. 
Overall format quality is the fraction of records with \emph{all} slots passing:
\[
\mathrm{FPR}_{\text{overall}} \;=\; \frac{1}{N}\sum_{n=1}^{N}\mathbf{1}\{\forall f\in\mathcal{F}:\ \text{lint\_pass}(x^{(n)}_f)\}.
\]

\paragraph{Overall style.}
Per (doc,slot) we compute an input-aware style score $s^{(n)}_f$:
\[
s^{(n)}_f=\begin{cases}
1,& \text{gold empty}\ \wedge\ \text{pred empty}\\
0,& \text{gold empty}\ \wedge\ \text{pred nonempty}\\
0,& \text{gold nonempty}\ \wedge\ \text{pred empty}\\
\cos\big(\phi(\mathrm{mask}(\text{pred})),\,\phi(\mathrm{mask}(\mathrm{exemplar}_f))\big),& \text{gold nonempty}\ \wedge\ \text{pred nonempty}
\end{cases}
\]
where $\mathrm{mask}$ maps letters$\!\to\!T$, digits$\!\to\!D$, keeps punctuation; $\phi$ is a normalized char-3gram TF vector; locations use the max over canonical forms and list slots treat each item as a single $T$. 
We report the macro style score:
\[
\mathrm{SC}_{\text{macro}} \;=\; \frac{1}{|\mathcal{F}|}\sum_{f\in\mathcal{F}}\left(\frac{1}{N}\sum_{n=1}^{N} s^{(n)}_f\right).
\]


\subsection{Latency and Cost at Inference (R6)}
\label{subsec:latency-cost}

Operational efficiency is measured per document from \texttt{timing.duration\_ms}, reporting p50 (median), p90, p95, p99, mean, min, and max. 
When available, we also estimate cost from API token usage (prompts, completions, and embeddings), reported as USD per document. 
Because retrieval and embedding are lightweight relative to decoding, schema- and example-guided approaches remain compatible with near real-time use.

\subsection{User Correction and Human-in-the-Loop Reliability (R4)}
\label{subsec:correction}

Although MUC-4 does not capture interactive editing, \textit{Invox} surfaces all auto-filled fields for review with typed widgets (date pickers, enum selectors, multi-value inputs). 
This design safely combines ``empty-when-uncertain'' behavior with rapid human confirmation, preserving downstream data quality.

\subsection{Modularity and Extensibility (R5)}
\label{subsec:modularity}

The architecture decouples retrieval, prompting, decoding, and verification into swappable modules. 
We evaluate alternative extraction strategies under the same metrics (OBS/NES/EAI, SF1/Chamfer, FDA/HR/MR/RFA, latency), enabling targeted optimization. 
Because slots are declarative (descriptions, types, allowed values), new domains or fields integrate without changing the evaluation stack.

\paragraph{Summary.}
Our protocol reports (i) \textbf{content similarity} (SoftCoverage, SoftSpecificity, Chamfer, SF1), (ii) \textbf{decision quality} (FDA, HR, MR, RFA) with an empty-field sensitivity index (EAI), under both \textbf{baseline} and \textbf{gold-nonempty-only} regimes, and (iii) \textbf{operational efficiency} (latency, cost). 
Together these satisfy requirements on accuracy (R2), consistency (R1), domain alignment (R3), modularity (R5), operational efficiency (R6), and human oversight (R4) in a single, coherent evaluation framework.
