This chapter presents the conceptual design of a privacy-preserving, retrieval-augmented vulnerability detection and repair system integrated into Visual Studio Code. The core objective is to assist developers in identifying and remediating security vulnerabilities in JavaScript and TypeScript code using locally deployed Large Language Models (LLMs), without transmitting source code to external services.

The design is motivated by three critical shortcomings observed in existing work. First, traditional Static Application Security Testing (SAST) tools rely on predefined rules and struggle with contextual reasoning, producing false positives when security-relevant patterns appear in benign contexts \cite{sheng2025survey,li2025iris}. Second, cloud-based LLM assistants require transmitting proprietary source code to external servers, which is unacceptable in regulated or security-sensitive environments \cite{kaur2025cyberreview,gholami2024llmcyber}. Third, current approaches provide limited explainability and repair guidance, making it difficult for developers to understand and act upon detected vulnerabilities \cite{johnson2013don,christakis2016developers}.

The proposed system addresses these limitations through a modular architecture that enforces separation of concerns while preserving end-to-end coherence. The design directly supports the six requirements established in Section~\ref{sec:requirements}: detection consistency is enforced through retrieval-augmented grounding (R1), contextual reasoning is enabled through code analysis and data flow understanding (R2), intermediate artifacts and explanations provide transparency (R3), concrete repair suggestions are generated based on security knowledge (R4), all processing occurs locally without external data transmission (R5), and latency is optimized for IDE-integrated workflows (R6).

The remainder of this chapter is organized as follows. Section~\ref{sec:concept-derivation} derives the conceptual design from the analysis results, explaining how each requirement motivates specific architectural decisions. Section~\ref{sec:system-components} describes the core components that comprise the system, detailing their responsibilities, inputs, outputs, and interactions. Section~\ref{sec:detection-strategies} presents detection and analysis workflows that illustrate how the system operates in different scenarios. Section~\ref{sec:system-architecture} provides the high-level system architecture and design, including context diagrams, component views, and process flows. Detailed implementation and experimental validation are deferred to Chapters~\ref{chap:implementation} and~\ref{chap:evaluation}, respectively.

\input{src/chapters/concept/concept_derivations}

\input{src/chapters/concept/components}
\input{src/chapters/concept/strategies}
\input{src/chapters/concept/system_architecture_and_design}